---
title: "Homework 6"
author: "Cynthia Liu (cl3938)"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(MASS)
library(modelr)
library(mgcv)
library(purrr)
library(rnoaa)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


### Problem 1

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  dplyr::select(city_state, resolution, victim_age, victim_race, victim_sex)
```


Start with one city.

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  dplyr::select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```


Try this across cities.

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  dplyr::select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  dplyr::select(city_state, term, OR, starts_with("CI")) 
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



### Problem 2

First, let's import and clean the data. The following steps were taken: 

  * Import CSV 
  * Clean names
  * Change relevant variables to factor
  
```{r}
#input data
baby_df = 
  read_csv("./data/birthweight.csv") %>%
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)) 
```
There are `r sum(is.na(baby_df))` missing values in the entire dataframe. This is great, so we'll progress to making our regression models.

Model 1: curated model

For this model, we'll use a stepwise algorithm to automatically select the best model from the full list of predictors. This is primarily used as exploratory analysis: the model adds and removes predictors iteratively based on a predetermined criteria. For my purposes, I'm going to use Akaike Information Criteria, or AIC, a metric that is a good balance between prioritizing model fit and also penalizing for too many predictors.

It's important to note that stepwise model selection results shouldn't be overintepreted, and there's a great deal of uncertainty surrounding them. I'm only going to use it to explore the data a bit, since I know next to nothing about this field - true models should always be based in reason & literature review.

Let's get started!
```{r}
#Fit the full model 
full_bwt_linear.model <- lm(bwt~., data = baby_df)

#Stepwise regression model
step_bwt_linear.model <- stepAIC(full_bwt_linear.model, direction = "both", 
                      trace = FALSE)

#Display converged model
step_bwt_linear.model %>% 
  broom::tidy() %>%
  knitr::kable(digits = 3)

#Plot the model
baby_df %>% 
  modelr::add_residuals(step_bwt_linear.model) %>% 
  modelr::add_predictions(step_bwt_linear.model) %>%
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  geom_smooth(se = F) +
  labs(
    x = "Predicted Values",
    y = "Residuals",
    title = "Residuals vs Fitted"
  )

```

Other models: 

* Model 1: length at birth and gestational age as predictors (main effects only)
```{r}
test_1.model <- lm(bwt~gaweeks + blength, data= baby_df)

test_1.model %>% 
  broom::tidy() %>%
  knitr::kable(digits = 3)
```

* Model 2: head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
test_2.model <- lm(bwt~babysex*bhead*blength, data = baby_df)

test_2.model %>% 
  broom::tidy() %>%
  knitr::kable(digits = 3)
```

Before we do cross-validation, let's take a look at goodness of fit of our 3 respective models:
```{r}
baby_df %>% 
  gather_predictions(step_bwt_linear.model, test_1.model, test_2.model) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

Let's start with cross-validation! Splitting data into training & testing:
```{r}
cv_df =
  crossv_mc(baby_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

Fitting models & obtaining RMSE values: 
```{r}
cv_df = 
  cv_df %>% 
  mutate(
    step_mod  = map(train, ~lm(bwt~babysex + bhead + blength + delwt + fincome + gaweeks + mrace + parity + ppwt + smoken, data = .x)),
    test1_mod = map(train, ~lm(bwt~gaweeks + blength, data = .x)),
    test2_mod = map(train, ~lm(bwt~babysex*bhead*blength, data = .x))) %>% 
  mutate(
    rmse_step = map2_dbl(step_mod, test, ~rmse(model = .x, data = .y)),
    rmse_test1  = map2_dbl(test1_mod, test, ~rmse(model = .x, data = .y)),
    rmse_test2 = map2_dbl(test2_mod, test, ~rmse(model = .x, data = .y)))
```

Okay, now let's make our violin plots:

```{r}
cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

### Problem 3

Import data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  dplyr::select(name, id, everything())
```

